---
title: "Fitting Normal Dists"
author: "Francis Windram"
date: "16/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fitdistrplus)
```

Make a simple normal distribution vector as a dummy

```{r}
speed <- 0.25
SD <- 0.125
logspeed <- log(speed)
logSD <- log(SD)

rautonorm <- function(n,mean=0,sd=1,r){
  ranfunc <- function(i,z,r) sqrt(1-r^2) * sum(z[2:(i+1)]*r^(i-(1:i))) + z[1]*r^i # for the autocorrelation
  # this is a known eqn that someone's worked out somewhere - he's lost the reference now though
  z <- rnorm(n)
  mean + sd*c(z[1], sapply(1:(n-1), ranfunc, z, r))
  }

testvec <- exp(rautonorm(1000, logspeed, logSD, 0.2))
print(mean(testvec))
```


```{r}
plotdist(testvec, histo = TRUE, demp = TRUE)
```

```{r}
plotdist(log(testvec), histo = TRUE, demp = TRUE)
```


```{r eval=FALSE, include=FALSE}
descdist(log(testvec), discrete=FALSE, boot=500)
```

None of this matters as we know our fits are normal in log space, sooooo

```{r}
testfit <- fitdist(log(testvec), "norm")
summary(testfit)
```

```{r}
print(paste0("Input Mean: ", exp(as.double(testfit$estimate[1]))))
print(paste0("Input SD: ", exp(0-as.double(testfit$estimate[2])))) # NOTE this is NOT the actual SD of the data in realspace
# print(paste0("Coefficient of Variation: ", exp(0-as.double(testfit$estimate[2])) / exp(as.double(testfit$estimate[1]))))
print(paste0("Coefficient of Variation in log space: ", as.double(0-testfit$estimate[2]) / as.double(testfit$estimate[1])))
```

*CAREFUL HERE! The +x SD == -x SD, thus it can be confusing to backtransform the SD to realspace!!! You MAY have to negate it as I do above*

Though let's just check how it deals with fitting lognormality to the og data

```{r}
testfit2 <- fitdist(testvec, "lnorm")
summary(testfit2)
```

Right... so we don't necessarily know whether we expect the CV in log space or the CV in linear space to be the one we want to see. 

One helper func would be something that derives the SD for a mean given a particular logCV.

*NOTE: the logCV is the CV of the fit distribution in log space. This is NOT the same as log(CV)*

```{r}
# Find exp(logSD) in realspace from mean in realspace and desired CV in logspace
# Note that exp(logSD) != SD, these are not the same thing!!!!! 
find_SD_from_logCV <- function(mean, logCV){
  return(exp(log(mean) * logCV))
}

# Find SD in realspace from mean in realspace and desired CV in realspace
find_SD_from_CV <- function(mean, CV){
  return(mean * CV)
}
```

**Note: the below does NOT come up with the SD of the original data!!!!**

```{r}
print(find_SD_from_logCV(0.05, 1.2))
```

```{r}
find_logCV_from_lognormal <- function(v, autocorrect_neg_SD = TRUE, suppress_warnings = TRUE){
  # This function finds the CV in logspace of a lognormally distributed vector.
  # Mathematically this isn't really a measurement, and it does not necessarily behave nicely
  # However the geometric CV is independent of the geometric mean...so we are a bit stuck
  # Args:
  # v = data
  # autocorrect_neg_SD = automatically negate SD if mu is negative
  #
  # Returns:
  # logCV: Coefficient of Variation in log space
  
  if (suppress_warnings){
    suppressWarnings(fit <- fitdist(v, "lnorm"))
  } else {
    fit <- fitdist(v, "lnorm")
  }
  
  mu <- as.double(fit$estimate[1])
  sigma <- as.double(fit$estimate[2])
  # Note: If mu is exactly 0, this will return a Inf.
  # It is safest to check for this using is.finite()
  
  if (mu < 0 & autocorrect_neg_SD){
    # If mu is negative, then a positive sigma is unlikely to be the case
    # as the scale in realspace would be orders of magnitude off.
    sigma <- 0 - sigma
  }
  
  return(c("meanlog"=mu, "meansd"=sigma, "logCV"=sigma / mu))
}
```

## Final test!

```{r}
find_logCV_from_lognormal(testvec)
find_SD_from_logCV(speed, as.double(find_logCV_from_lognormal(testvec)[3]))
```

```{r eval=FALSE, include=FALSE}
speeds <- 10^seq(1,5)
SDs <- speeds / 5
CVs <- SDs / speeds
logCVs <- log(SDs) / log(speeds)

print(CVs)
print(logCVs)
```

## Final note

So after some more hefty googling, it looks like logCV is a little shakey as far as a measure of dispersion when the logged data is negative (and thus the logmean and logSD are negative). HOWEVER as a method to get a general idea for reasonable parameters to input, it might be better than calculating the CV from $e^{\sigma}$ of a $\sigma$ calculated in log space. Essentially the SD parameter we are using at the top is NOT representative of the SD in realspace, it is instead the aforementioned problematic $e^{\sigma}$. ~~Basically if we provide logSD to `pathgen()` as `log(SD)`, then that fundamental `SD` parameter does not mean anything itself, and has no biological meaning.~~

EDIT: So it turns out that $e^{stdev(\ln{v})}$ is actually termed the [geometric standard deviation](https://en.wikipedia.org/wiki/Geometric_standard_deviation). This is a measure of log-normal dispersion related to the geometric mean, analogous to the relationship between mean and SD for normally distributed data. There is a geometric coefficient of variation, but frustratingly it is entirely independent of the geometric mean.

Long story short, if we calculate logCVs for a bunch of original speed distributions it MIGHT be lucky enough to give us a consistent logCV that we can use to parameterise input SDs given desired speeds. It MIGHT also give us some horrible relationship that we'd need to model in order to find a consistent relationship between the two parameters (logMean and logSD). This is messy stuff, and I think we'll have to try it out on real data to see if it might work.

Alternatively I found one example of someone calculating a "coefficient of variation" from the GSD/GM, but literally only one: https://www.mdpi.com/2075-163X/9/3/193/htm

I really don't know if that is something that is mathematically valid or useful for our approach but hey.
