---
title: "Triangle_Sim"
author: "Francis Windram"
date: "06/08/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
A <- c(0,10)
B <- c(10,10)
C <- c(5,0)

trix <- c(A[1], B[1], C[1])
triy <- c(A[2], B[2], C[2])

l = c(1/3, 1/3, 1/3)

x = sum(trix * l)
y = sum(triy * l)
print(x)
print(y)
```

```{r}
plot(trix, triy)
polygon(trix, triy, border = "green")
points(x, y)
```

Generate triangle with given view angle and effective distance

```{r}
theta = pi/8
dist = 20
origin = c(0,0)
C = origin

xshift = tan(theta) * dist

A = c(origin[1] - xshift, dist)
B = c(origin[1] + xshift, dist)
C = origin

trix <- c(A[1], B[1], C[1])
triy <- c(A[2], B[2], C[2])

# tan = opp / adj
# opp = tan * adj
```


Draw a bunch of lines between two edges of the triangle.

(We do this by defining points in the triangle in a barycentric coordinate system and then converting them to cartesian space)
```{r}

shifter <- function(x, n = 1) {
  if (n == 0) x else c(tail(x, -n), head(x, n))
}

plot(trix, triy)
polygon(trix, triy, border = "green")

n = 100000

lengthvec = rep(0, n)

for (i in 1:n){
  bc1_1 <- runif(1, min = 0, max = 1)
  bc1_2 <- 1-bc1_1
  bc1 = c(bc1_1, bc1_2, 0)
  
  bc2_1 <- runif(1, min = 0, max = 1)
  bc2_2 <- 1-bc2_1
  bc2 = c(bc2_1, bc2_2, 0)

  shift = zero_idxs <- sample(0:2, 2, replace = FALSE)

  bc1 <- shifter(bc1, zero_idxs[1])
  bc2 <- shifter(bc2, zero_idxs[2])

  xs = c(sum(trix * bc1), sum(trix * bc2))
  ys = c(sum(triy * bc1), sum(triy * bc2))

  lines(xs, ys, col = "red")
  lengthvec[i] <- sqrt((xs[2] - xs[1])**2 + (ys[2] - ys[1])**2)
}
df <- data.frame(lengthvals = lengthvec)
```

Plot the CDF

At a given $x$ value, the y value is the probability of the length of a line across the triangle being less than or equal to $x$

```{r}
library(ggplot2)
p <- ggplot(df, aes(x=lengthvals)) + stat_ecdf()# + geom_density()
print(p)
```


Plot the PDF for good measure

```{r}
p <- ggplot(df, aes(x=lengthvals)) + geom_density()
print(p)
```